---
title: "02_statistical_learning"
author: "Julen Pardo"
date: "4/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Applied exercises

## 8

### 8.a

```{r}
college = read.csv("../datasets/College.csv")
```

### 8.b

```{r}
rownames(college) = college[,1]
college = college[,-1]
college

```

### 8.c (i)

```{r}
summary(college)
```
### 8.c (ii)
```{r}
pairs(college[,2:10])
```

### 8.c (iii)
```{r}
boxplot(Outstate~Private, data=college)
```

### 8.c (iv)
```{r}
Elite = rep("No", nrow(college))
Elite[college$Top10perc > 50] = "Yes"
Elite = as.factor(Elite)
college = data.frame(college, Elite)

summary(college)

boxplot(Outstate~Elite, data=college)
```

### 8.c (v)

```{r}
par(mfrow=c(2, 2))
hist(college$Apps, breaks=20, xlim=c(0, 20000), ylim=c(0, 500))
hist(college$Enroll, xlim=c(0, 4000))
hist(college$PhD, breaks=20)
hist(college$Top10perc)
```

## 9

```{r}
auto = read.csv("../datasets/Auto.csv")
auto
```
### 9.a

*Quantitative*

* mpg
* displacement
* horsepower
* weight
* acceleration
* year

*Qualitative*

* year
* origin

### 9.b

```{r}
sapply(auto[,1:7], range)
```

### 9.c

```{r}
sapply(auto[,1:7], mean)
sapply(auto[,1:7], sd)
```

### 9.d

```{r}
auto_subset = auto[-c(10:84),]

sapply(auto_subset[,1:7], range)
sapply(auto_subset[,1:7], mean)
sapply(auto_subset[,1:7], sd)
```

### 9.e

```{r}
pairs(auto[,1:7])
```

*Positive linear correlations*

* Weight and displacement
* Weight and horsepower

*Negative, non linear, monotonic correlations*

* mpg and displacement
* mpg and horsepower
* mpg and weight

### 9.f

* The greater the displacement, horsepower or weight is, the lower the mpg


## 10

### 10.a

```{r}
library(MASS)
dim(Boston)
```

### 10.b

```{r}
pairs(~crim + age + dis + medv + rm + nox, data=Boston)
```

### 10.c

```{r}
cor(Boston[-1], Boston$crim)
```

### 10.d

```{r}
par(mfrow=c(1, 2))
boxplot(Boston$crim)
hist(Boston$crim, xlim=c(0, 40), breaks=40)
summary(Boston$crim)
sd(Boston$crim)
third_quartile = quantile(Boston$crim, 0.75)
iqr = IQR(Boston$crim)
Boston[Boston$crim > third_quartile + 1.5 * iqr,]
```

The crime distribution is far from normal and there are several outliers that
skew the mean (mean 3,61 vs 0.25 median, being the standard deviation 8.6).
There are 66 suburbs with a crime rate greater than the upper whisker in the
box plot. The range is quite wide but the majority of the occurrences belong
to one edge.

### 10.e

```{r}
nrow(Boston[Boston$chas == 1,])
```
### 10.f

```{r}
median(Boston$ptratio)
```

### 10.g

```{r}
row.names(Boston[Boston$medv == min(Boston$medv),])
```

### 10.h

```{r}
nrow(Boston[Boston$rm > 7,])
```

```{r}
nrow(Boston[Boston$rm > 8,])
```

```{r}
as.data.frame(apply(Boston, 2, summary))
as.data.frame(apply(Boston[Boston$rm > 8,], 2, summary))
```

Lower crime and lstat, much higher medv